A Quantitative Genetic Linked Selection Model fit to the Negative Selection Process in Humans

### 2023-07-05 manuscript/manuscript.tex
ms changes

\subsection*{Estimates of Negative Selection Parameters in Humans}

@@

However, across all of our annotation models, new deleterious mutations in
putatively conserved features are consistently estimated to have strongly
deleterious effects (Figure \ref{fig:figure-3}A). This includes features such
as coding sequences, PhastCons regions, or the top 6\% of basepairs with the
most pathogenic CADD scores. Thus, even with a model that can better
distinguish weak versus strong selection, our DFE estimates suggest

qualitatively
agree with recent work \parencite{Murphy2022-sj}.

Moreover, the estimated mutation rates are close
to pedigree-based estimates for most of these annotation models (XXX).

Second, we find evidence that our method will indeed estimate weak selection
across features

Given that the B' method more
accurately predicts diversity across weaker selection coefficients, we wondered
whether our inferred DFEs could

Linked selection theory suggests the reduction signal is determined by the
spatial distributions of recombination and fitness variance along the
chromosome. When fitness variation is determined mostly by negative selection,
the

Where does the information for the DFE come from?

All of our models show a general trade off in fitting the linked selection
signal.

Next, we developed a composite likelihood method to estimate expected diversity
without linked selection $\pi_0$, the deleterious mutation rate ($\mu_D$), and
distribution of fitness effects ($\mathbf{W}$) under negative selection.
Because the focus of our work is to fix deficiencies in the classic BGS model,
our current implementation does not consider reductions in diversity due to
recurrent hitchhiking or other positive selection processes. Since previous
work has established that purifying selection is the dominant mode of linked
selection in humans and hitchhiking plays a limited role
\parencite{McVicker2009-ax,Murphy2022-sj}, humans are a suitable organism to
test our method.

@@

Given that previous methods used classic BGS theory to fit genome-wide patterns
of human diversity, we were curious whether their qualitative findings would
change once weak selection was more accurately predicted. In particular, the
U-shaped relationship between $B'$ and the strength of selection suggests that
a given reduction factor could be explained by two different selection
coefficients.

@@

>>>>

\subsection*{Estimates of the Distribution of Fitness Effects in Humans}

@@

However, across all of our annotation models, new deleterious mutations in
likely conserved features are consistently estimated to have strongly
deleterious effects (Figure \ref{fig:figure-3}A) consistent with previous work
\parencite{McVicker2009-ax,Murphy2022-sj}. This includes features such as
coding sequences, PhastCons regions, or the top 6\% of basepairs with the most
pathogenic CADD scores. However, we find that there is considerable uncertainty
in how strong selection is against the most deleterious mutations. For example,
in the CADD 6\% full-track model, new mutations are estimated to have a
selection coefficients of $s = 10^{-1}$ with 42\% chance, $s=10^{-2}$ with
7.8\% chance, and $s=10^{-3}$ with 41\% chance. These estimates suggest a
bimodal DFE (Figure \ref{fig:figure-3}A, second row, left column), and are
mirrored across the other CADD models.

Given that previous methods have typically not estimated the DFE up to
selection coefficients of $s=10^{-1}$, we ran additional models using a DFE
constrained to $s=10^{-2}$. Each of these constrained-DFE models across all
populations fit the data less well than models run with the grid up to
$s=10^{-1}$, explaining 1\% less of the leave-one-chromosome-out variation on
average. As expected, the constrained-DFE models put more mass on $s=10^{-2}$
(see Supplementary Table \ref{supp:tbl-r2}). Overall, this common behavior
across models seems likely to reflect a non-identifiability issue between
selective effects of $s=10^{-1}$ and $s=10^{-2}$, which we refer to as the
\emph{strong-selection non-identifiability} issue.

We propose that one mechanism for this non-identifiability is that mutations
extremely harmful to fitness are difficult to distinguish from a smaller
effective population. Intuitively, as $s$ increases, the average persistence
time (in generations) of the mutation in the population goes below one
generation, i.e. death of the carrier of the \emph{de novo} mutation. In this
hypothetical limit, each occurrence of a very harmful mutation is like an
instantaneous death and thus acts like a genome-wide reduced effective
population size. In our model, this genome-wide effective population size is
parameterized by $\pi_0$; thus, one line of evidence for this effect is a
correlation between estimates of $\pi_0$ and average selection $\bar{s}$.
Looking at the CADD 6\% models fit across the default and constrained DFE
grids, we find this is the case. Across every population, average selection was
about 7 times larger using the default grid, and $\pi_0$ was $5.6\%$ larger.
However, there was no similar consistent change in mutation rate estimates
among populations. Overall, this suggests that the linked selection signal
alone carries little information that can differentiate very strong selection
from a slightly smaller drift-effective population size.

Despite the strong-selection non-identifiability issue, we still were curious
how stable these estimates are across samples from reference populations. In
particular, we focused on the Feature Priority model, which fits the data less
well but is more interpretable since its annotation is based on gene features.
Overall, estimates of the distribution of selection coefficients are stable
across populations and had narrow uncertainty as estimated by a block jackknife
procedure (Figure \ref{figure-3}B). Coding sequences show the strong-selection
non-identifiability issue, and an uncertain signal of a neutral mode in the
DFE. This would be expected given the very different levels of selective
constraint synonymous and non-synonymous sites in coding sequences are under.
The PhastCons class of this Feature Priority model, which contains non-coding
highly conserved elements (XXX), is under strong selective constraint as
expected.

\subsection*{Estimates of the Deleterious Mutation Rate in Humans}

@@

@@

- the $pi_0$ problem
